% Copyright 2018 Beckman Coulter, Inc.
%
% Permission is hereby granted, free of charge, to any person
% obtaining a copy of this software and associated documentation files
% (the "Software"), to deal in the Software without restriction,
% including without limitation the rights to use, copy, modify, merge,
% publish, distribute, sublicense, and/or sell copies of the Software,
% and to permit persons to whom the Software is furnished to do so,
% subject to the following conditions:
%
% The above copyright notice and this permission notice shall be
% included in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
% EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
% MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND
% NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS
% BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN
% ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
% CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\lstdefinelanguage{json}{
  basicstyle=\upshape\mdseries\frenchspacing\ttfamily,
  columns=flexible,
  string=[s]{"}{"},
  stringstyle=\color{blue},
  comment=[l]{:},
  commentstyle=\color{black},
  escapeinside={|}{|},
  frame=single
}

\chapter {JSON Interface}

\section {Introduction}

The programming interface provides support for JavaScript Object
Notation (JSON)~\cite{RFC7159}.

This implementation defines a mapping between JSON values and Scheme
values as follows:

\begin{center}
\begin{tabular}{ll}
  JSON & Scheme \\ \hline

  \code{true}  & \code{\#t} \\
  \code{false} & \code{\#f} \\
  \code{null}  & symbol \code{null} \\
  string       & string \\
  number       & number \\
  array        & list \\
  object       & hashtable mapping symbols to values \\

  \hline
\end{tabular}
\end{center}

JSON objects are represented as hashtables whose keys are symbols
derived from the corresponding JSON object member names.

No validation is performed to ensure that numeric values fall within
the range supported by typical JavaScript implementations. As a result,
some values produced by this implementation may not be portable to all
JSON consumers.

By default, this implementation accepts input conforming to
JSON5~\cite{JSON5}, a superset of JSON. This allows additional syntax
not permitted by standard JSON, including comments, single-quoted
strings, and numeric values such as \code{Infinity}, \code{-Infinity},
and \code{NaN}.

Output is, by default, written in standard JSON format. However, the
implementation permits the non-standard numeric values
\code{Infinity}, \code{-Infinity}, and \code{NaN} to be written as
well. This behavior is controlled by an option that is enabled by
default.

\section {Programming Interface}

\defineentry{json:extend-object}
\begin{syntax}
  \code{(json:extend-object \var{ht} [\var{key} \var{value}] \etc)}
\end{syntax}

The \code{json:extend-object} construct adds the \var{key} /
\var{value} pairs to the hashtable \var{ht} using
\code{hashtable-set!}. Each \var{key} is a literal identifier
or an unquoted expression \code{,\var{e}} that evaluates to a
symbol. The resulting expression returns \var{ht}.

\defineentry{json:make-object}
\begin{syntax}
  \code{(json:make-object [\var{key} \var{value}] \etc)}
\end{syntax}

The \code{json:make-object} construct expands into a call to
\code{json:extend-object} with a new hashtable.

\defineentry{json:object?}
\begin{procedure}
  \code{(json:object? \var{x})}
\end{procedure}
\returns{} a boolean

The \code{json:object?} procedure determines whether or not the datum
\var{x} is an object created by \code{json:make-object}.

\defineentry{json:cells}
\begin{procedure}
  \code{(json:cells \var{ht})}
\end{procedure}
\returns{} a vector

The \code{json:cells} procedure returns a vector containing the
cells of the underlying hashtable.

\defineentry{json:size}
\begin{procedure}
  \code{(json:size \var{ht})}
\end{procedure}
\returns{} an integer

The \code{json:size} procedure returns the number of cells
in the underlying hashtable.

\defineentry{json:delete"!}
\begin{procedure}
  \code{(json:delete! \var{ht} \var{path})}
\end{procedure}
\returns{} unspecified

The \code{json:delete!} procedure expects \var{path} to be a symbol or
a non-empty list of symbols.
If \var{path} is a symbol, then \code{json:delete!} is equivalent
to \code{hashtable-delete!}.
Otherwise, \code{json:delete!} follows \var{path} as it descends into
the nested hashtable \var{ht}, treating each element as a key into
the hashtable reached by traversing the preceding elements.
When \code{json:delete!} reaches the final key in \var{path},
it calls \code{hashtable-delete!} to remove the association for
that key in the hashtable reached at that point.
If any key along the way does not map to a hashtable,
\code{json:delete!} has no effect.

\defineentry{json:ref}
\begin{procedure}
  \code{(json:ref \var{ht} \var{path} \var{default})}
\end{procedure}
\returns{} the value found by traversing \var{path} in \var{ht},
\var{default} if none

The \code{json:ref} procedure expects \var{path} to be a symbol or
a non-empty list of symbols.
If \var{path} is a symbol, then \code{json:ref} is equivalent
to \code{hashtable-ref}.
Otherwise, \code{json:ref} follows \var{path} as it descends into
the nested hashtable \var{ht}, treating each element as a key into
the hashtable reached by traversing the preceding elements.
When \code{json:ref} reaches the final key in \var{path},
it calls \code{hashtable-ref} to retrieve the value of that key
in the hashtable reached at that point.
If any key along the way does not map to a hashtable,
or if the final hashtable does not contain the final key,
\code{json:ref} returns \var{default}.

\defineentry{json:set"!}
\begin{procedure}
  \code{(json:set! \var{ht} \var{path} \var{value})}
\end{procedure}
\returns{} unspecified

The \code{json:set!} procedure expects \var{path} to be a symbol or
a non-empty list of symbols.
If \var{path} is a symbol, then \code{json:set!} is equivalent
to \code{hashtable-set!}.
Otherwise, \code{json:set!} follows \var{path} as it descends into
the nested hashtable \var{ht}, treating each element as a key into
the hashtable reached by traversing the preceding elements.
When \code{json:set!} reaches the final key in \var{path},
it calls \code{hashtable-set!} to set that key in the
hashtable reached at that point.
If any key along the way does not map to a hashtable,
\code{json:set!} installs an empty hashtable at that key
before proceding.
If \var{path} is malformed at some point, \code{json:set!} may
still mutate hashtables along the valid portion of the path
before reporting an error.

\defineentry{json:update"!}
\begin{procedure}
  \code{(json:update! \var{ht} \var{path} \var{procedure} \var{default})}
\end{procedure}
\returns{} unspecified

The \code{json:update!} procedure expects \var{path} to be a symbol or
a non-empty list of symbols.
If \var{path} is a symbol, then \code{json:update!} is equivalent
to \code{hashtable-update!}.
Otherwise, \code{json:update!} follows \var{path} as it descends into
the nested hashtable \var{ht}, treating each element as a key into
the hashtable reached by traversing the preceding elements.
When \code{json:update!} reaches the final key in \var{path},
it calls \code{hashtable-update!} to update that key in the
hashtable reached at that point.
If any key along the way does not map to a hashtable,
\code{json:update!} installs an empty hashtable at that key
before proceding.
If \var{path} is malformed at some point, \code{json:update!} may
still mutate hashtables along the valid portion of the path
before reporting an error.

\defineentry{json:read}
\defineentry{json:read-options}
\phantomsection % make pageref go to correct page for this label
\label{json:read-options}
\begin{procedure}
  \code{(json:read \var{ip} \opt{\var{options}})}
\end{procedure}
\returns{} a Scheme object or the eof object

The \code{json:read} procedure reads characters from the textual input
port \var{ip} and returns an appropriate Scheme object. If no more
input is available, the end-of-file object is returned.

The optional \var{options} argument is constructed using
\code{(json:read-options (\var{option} \var{value}) \dots)}. The
following options are supported:

\begin{tabular}{lrp{.5\textwidth}}
  option & default & description \\ \hline

  \code{extended-identifiers?} & \code{\#f} &
  If true, enables more permissive reading of unquoted symbols. In
  value position, this applies to both JSON and JSON5 input. If
  \code{json5?} is enabled, it also applies to object member
  names. When enabled, unquoted symbols may include characters not
  permitted by the JSON5 identifier syntax (for example, hyphenated
  identifiers in the style of Scheme). If false, unquoted symbols must
  conform to the JSON5 identifier syntax. \\

  \code{inflate-object} & \code{\#f} &
  When a JSON object is read, it is first converted into a hashtable.
  If \code{inflate-object} is provided, it is then applied to the
  hashtable to perform application-specific transformation. By
  default, no transformation is performed. \\

  \code{inflate-symbol} & \code{\#f} &
  Specifies a procedure of one argument for interpreting unquoted
  symbols in value position. The procedure is called with the symbol
  as its argument and may return any Scheme object. If
  \code{inflate-symbol} is \code{\#f}, the occurrence of an unquoted
  symbol in value position is an error. This behavior does not apply
  to object member names. \\

  \code{json5?} & \code{\#t} & If \code{\#t}, input is parsed according
  to JSON5. If \code{\#f}, only standard JSON syntax is accepted.

\end{tabular}

The following exceptions may be raised:
\begin{itemize}
\item \code{invalid-surrogate-pair}
\item \code{\#(json:unexpected-input \var{context} \var{data} \var{input-position} \var{name})}
\end{itemize}

\defineentry{json:write}
\defineentry{json:write-options}
\phantomsection % make pageref go to correct page for this label
\label{json:write-options}
\begin{procedure}
  \code{(json:write \var{op} \var{x} \opt{\var{indent}} \opt{\var{options}})}
\end{procedure}
\returns{} unspecified

The \code{json:write} procedure writes the object \var{x} to the
textual output port \var{op} in JSON format. By default, \code{json:write}
sorts the keys of JSON objects using \code{string<?}, after converting
the symbolic keys to strings.
This sort order provides stable output.
The \code{json:key<?} parameter may be used to alter or disable key sorting.
Scheme fixnums, bignums, and finite flonums may
be used as numbers.

When \var{indent} is a non-negative fixnum, the output is more
readable by a human. List items and key/value pairs are indented on
individual lines by the specified number of spaces. When \var{indent}
is 0, a newline is added to the end of the output. The default indent
of \code{\#f} produces compact output.

The \var{options} are defined with \code{(json:write-options
  (\var{option} \var{value}) \etc)}. The following options may be
used:

\begin{tabular}{lrp{.5\textwidth}}
  option & default & description \\ \hline

  \code{custom-write} & \code{\#f} & The \var{custom-write} procedure
  is invoked for any Scheme object that does not have a direct JSON
  representation (i.e., objects other than booleans, the symbol
  \code{null}, strings, and numbers). It may intervene to handle lists
  and hashtables differently or to handle objects that have no direct
  JSON counterpart.

  If \var{custom-write} does not handle a given object, it should
  return \code{\#f} to allow \code{json:write} to proceed with its
  default behavior.

  The \var{custom-write} procedure is called with four arguments: the
  textual output port \var{op}, the Scheme object \var{x}, the current
  \var{indent} level, and a writer procedure \var{wr} that should be
  used to write arbitrary Scheme objects. The \var{wr} procedure is
  equivalent to
  \code{(lambda (\var{op} \var{x} \var{indent})
    (json:write \var{op} \var{x} \var{indent} (json:write-options
      [custom-write \var{custom-write}])))}.

  The default value of \var{custom-write} is \code{\#f}, which
  indicates that the \code{json:custom-write} process parameter is
  used. \\

  \code{json5?} & \code{\#t} &
  If true, enables writing of the non-standard numeric values
  \code{Infinity}, \code{-Infinity}, and \code{NaN}. If false, output
  is restricted to standard JSON to maximize compatibility with other
  systems, and an exception is raised if any of these values are
  encountered.

\end{tabular}

If an object cannot be formatted, \code{\#(json:invalid-datum \var{x})}
is raised. The code does not check for cyclic data.

\defineentry{json:write-object}
\begin{syntax}
  \code{(json:write-object \var{op} \var{indent} \var{wr} [\var{key} \var{value}] \etc)}
\end{syntax}
\returns{} \code{\#t}

Given a textual output port \var{op}, an \var{indent} level, and a
writer procedure \var{wr}, the \code{json:write-object} construct
writes a JSON object with the given \var{key} / \var{value} pairs to
\var{op}. By default, \code{json:write-object} sorts the keys using
\code{string<?} on the string values of the symbols.
To alter or disable key sorting, see the \code{json:key<?} parameter.

Each \var{key} must be a distinct symbol.  The \var{wr}
procedure takes \var{op}, an object \var{x}, and an \var{indent} level
just like the \var{wr} procedure that is passed to \code{json:write}'s
\var{custom-write} procedure.

The following are equivalent, provided the keys are symbols and \code{json:key<?} has its default value.

\antipar\begin{alltt}
(begin (json:write \var{op} (json:make-object [\var{key} \var{value}] \etc) \var{indent}) \#t)
(json:write-object \var{op} \var{indent} json:write [\var{key} \var{value}] \etc)\end{alltt}\antipar

The latter trades code size and compile time for run-time efficiency.
At compile time, \code{json:write-object} sorts the keys and
preformats the strings that will separate values.

\defineentry{json:pretty}
\begin{procedure}
  \code{(json:pretty \var{x} \opt{\var{op}} \opt{\var{options}})}
\end{procedure}
\returns{} unspecified

The \code{json:pretty} procedure formats an object \var{x} to the
optional textual output port \var{op} in a human-readable form. The
object is indented and keys are sorted using
\code{natural-string-ci<?}. The default value of \var{op} is
\code{(current-output-port)}.

The optional \var{options} argument is defined using
\code{json:write-options} described in
Section~\ref{json:write-options}.

Calling \code{(json:pretty x)} is roughly equivalent to the following:
\antipar\codebegin
(parameterize ([json:key<? natural-string-ci<?])
  (json:write (current-output-port) x 0))
\codeend

\defineentry{json:custom-write}
\begin{parameter}
  \code{json:custom-write}
\end{parameter}
\returns{} a procedure or \code{\#f}

The \code{json:custom-write} process-parameter determines the default
\var{custom-write} argument for \code{json:write} and the other JSON
output procedures.

\defineentry{json:key<?}
\begin{parameter}
  \code{json:key<?}
\end{parameter}
\returns{} a boolean or a predicate for comparing two strings

The \code{json:key<?} parameter determines whether and how
\code{json:pretty}, \code{json:write}, and \code{json:write-object} sort keys
when writing JSON objects.
If set to \code{\#f}, keys are not sorted and the output order is arbitrary.
Otherwise, the output is sorted by obtaining a string for each key
via \code{symbol->string} and comparing those strings using the predicate
specified by \code{json:key<?}.
If \code{json:key<?} is set to \code{\#t}, which is the default,
then \code{json:pretty} sorts with \code{natural-string-ci<?}
while \code{json:write} and \code{json:write-object} sort with
\code{string<?}.
To affect the output of \code{json:write-object}, \code{json:key<?} must
be set to the desired value at expand time.

\defineentry{json:object->bytevector}
\begin{procedure}
  \code{(json:object->bytevector \var{x} \opt{\var{indent}} \opt{\var{options}})}
\end{procedure}
\returns{} a bytevector

The \code{json:object->bytevector} procedure calls \code{json:write}
on \var{x} with the optional \var{indent} and \var{options}, if
any, using a bytevector output port transcoded using
\code{(make-utf8-transcoder)} and returns the resulting bytevector.

The optional \var{options} argument is defined using
\code{json:write-options} described in
Section~\ref{json:write-options}.

\defineentry{json:bytevector->object}
\begin{procedure}
  \code{(json:bytevector->object \var{x} \opt{\var{options}})}
\end{procedure}
\returns{} a Scheme object

The \code{json:bytevector->object} procedure creates a bytevector input port
on \var{x}, calls \code{json:read} with the optional \var{options},
if any, and returns the resulting Scheme object after making sure the rest
of the bytevector is only whitespace.

The optional \var{options} argument is defined using
\code{json:read-options} described in Section~\ref{json:read-options}.

\defineentry{json:object->string}
\begin{procedure}
  \code{(json:object->string \var{x} \opt{\var{indent}} \opt{\var{options}})}
\end{procedure}
\returns{} a JSON formatted string

The \code{json:object->string} procedure creates a string output port,
calls \code{json:write} on \var{x} with the optional \var{indent} and
\var{options}, if any, and returns the resulting string.

The optional \var{options} argument is defined using
\code{json:write-options} described in
Section~\ref{json:write-options}.

\defineentry{json:string->object}
\begin{procedure}
  \code{(json:string->object \var{x} \opt{\var{options}})}
\end{procedure}
\returns{} a Scheme object

The \code{json:string->object} procedure creates a string input port
on \var{x}, calls \code{json:read} with the optional \var{options},
if any, and returns the resulting Scheme object after making sure the rest
of the string is only whitespace.

The optional \var{options} argument is defined using
\code{json:read-options} described in Section~\ref{json:read-options}.

\defineentry{json:write-structural-char}
\begin{procedure}
  \code{(json:write-structural-char \var{x} \var{indent} \var{op})}
\end{procedure}
\returns{} the new indent level

The \code{json:write-structural-char} procedure writes the character
\var{x} at an appropriate \var{indent} level to the textual output
port \var{op}. The character should be one of the following JSON
structural characters: \code{[ ] \{ \} : ,}

\defineentry{stack->json}
\begin{procedure}
  \code{(stack->json \var{k} \opt{\var{max-depth}})}
\end{procedure}
\returns{} a JSON object

The \code{stack->json} procedure renders the stack of continuation \var{k}
as a JSON object by calling \hyperlink{walk-stack}{\code{walk-stack}}.
The return value may contain the following keys:

\begin{tabular}{lp{4.6in}}
  \code{type} & \code{"stack"} \\
  \code{depth} & the depth of the stack \\
  \code{truncated} & if present, the \var{max-depth} at which the stack dump was truncated \\
  \code{frames} & if present, a list of JSON objects representing stack frames
\end{tabular}

A stack frame may contain the following keys:

\begin{tabular}{lp{4.6in}}
  \code{type} & \code{"stack-frame"} \\
  \code{depth} & the depth of this frame \\
  \code{source} & if present, a source object for the return point \\
  \code{procedure-source} & if present, a source object for the procedure containing the return point \\
  \code{free} & if present, a list of JSON objects representing free variables
\end{tabular}

A source object \var{x} with source file descriptor \var{sfd} is
represented by a JSON object containing the following keys:

\begin{tabular}{lp{4.6in}}
  \code{bfp} & \code{(source-object-bfp \var{x})} \\
  \code{efp} & \code{(source-object-efp \var{x})} \\
  \code{path} & \code{(source-file-descriptor-path \var{sfd})} \\
  \code{checksum} & \code{(source-file-descriptor-checksum \var{sfd})}
\end{tabular}

A free variable with value \var{val} is represented by a JSON object
containing the following keys:

\begin{tabular}{lp{4.6in}}
  \code{name} & a string containing the variable name or its index \\
  \code{value} & the result of \code{\fixtilde(format "~s" \var{val})} \\
\end{tabular}

\defineentry{json-stack->string}
\begin{procedure}
  \code{(json-stack->string \opt{\var{op}} \var{x})}
\end{procedure}
\returns{} see below

The two-argument form of \code{json-stack->string} prints the
stack represented by JSON object \var{x} to the textual output port \var{op}.
The single-argument form of \code{json-stack->string} prints the stack
represented by JSON object \var{x} to a string output port and returns
the resulting string.
In either case, the printed form resembles that generated by \code{dump-stack}
except that source locations are given as file offsets rather than line and character
numbers.
