% Copyright 2025 Indigo BioAutomation, Inc.
%
% Permission is hereby granted, free of charge, to any person
% obtaining a copy of this software and associated documentation files
% (the "Software"), to deal in the Software without restriction,
% including without limitation the rights to use, copy, modify, merge,
% publish, distribute, sublicense, and/or sell copies of the Software,
% and to permit persons to whom the Software is furnished to do so,
% subject to the following conditions:
%
% The above copyright notice and this permission notice shall be
% included in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
% EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
% MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND
% NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS
% BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN
% ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
% CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\chapter {Options Library}\label{chap:options}

\section {Introduction}

The \code{(swish options)} library provides a way to define a set of named options. The
options are stored in a record.

\section {Programming Interface}

\defineentry{define-options}
\begin{syntax}\begin{alltt}
(define-options \var{opt-name}
  (required [\var{key} \nt{req-spec} \etc] \etc)
  (optional [\var{key} \nt{opt-spec} \etc] \etc))\end{alltt}
\begin{grammar}
  \prodrow{\nt{req-spec}}{\code{(must-be \var{pred?} \etc)}}
  \altrow{\code{(filter \var{filter-expr})}}
  \prodrow{\nt{opt-spec}}{\nt{req-spec}}
  \altrow{\code{(default \var{default-expr})}}
\end{grammar}
\end{syntax}
\returns{} an \var{opt-name} macro and an \code{<\var{opt-name}>} record type that can be
used in \code{match} patterns

The \code{define-options} form generates an internal make-record procedure that evaluates
the \var{default-expr}, \var{pred?}, and \var{filter-expr} expressions left-to-right in
the order that fields appear in the definition. The order of evaluation for individual
expressions of a particular field is unspecified.  This means that default expressions are
generative. If this is not desirable, bind \var{default-expr} to a variable.

\var{opt-name} is an identifier. The \var{key}s are distinct identifiers.

The \var{opt-name} macro supports
\begin{itemize}

\item record construction and copying\begin{alltt}
(\var{opt-name} [\var{key} \var{val}] \etc)
(\var{opt-name} copy \var{obj} [\var{key} \var{val}] \etc)
(\var{opt-name} copy* \var{obj} [\var{key} \var{val}] \etc)\end{alltt}

\item record field access\begin{alltt}
((\var{opt-name} \var{key}) \var{obj})
(\var{opt-name} \var{key} \var{obj})\end{alltt}
    
\item record predicate\begin{alltt}
((\var{opt-name} is?) \var{obj})
(\var{opt-name} is? \var{obj})\end{alltt}

\end{itemize}
